\documentclass[11pt]{article}

% Minimal, stable packages; no external refs or urls
\usepackage[T1]{fontenc}
\usepackage{amsmath,amssymb}
\usepackage{mathtools}

\title{Recognition Architecture (Integrated):\\
Bridge $\rightarrow$ Constants $\rightarrow$ Masses $\rightarrow$ Gravity}
\author{Jonathan Washburn\\
Recognition Science, Recognition Physics Institute\\
Austin, Texas, USA\\
\texttt{jon@recognitionphysics.org}}
\date{} % keep blank

\begin{document}
\maketitle

\begin{abstract}
We present a single, falsifiable framework linking a proof--first, dimensionless layer to SI displays and to auditable predictions for particle spectra and cosmic dynamics. The layer order is: (i) a unique symmetric cost on the log axis, a quantized tick, the eight--tick minimal cycle in three--bit parity space, and the golden--ratio gap; (ii) a fixed ``Reality Bridge'' semantics with two independent SI landings and a single pass/fail inequality; (iii) a constants layer with a universal coherence quantum $E_{\mathrm{coh}}=\varphi^{-5}\,\mathrm{eV}$ and a common RG anchor $\mu_\star$; (iv) a finite constructor (ribbons $\&$ braids) that yields integer invariants and a closed--form residue map at $\mu_\star$, producing Standard--Model masses by fixed points without per--species knobs; (v) an information--limited gravity kernel used from galaxy rotation curves up to linear structure formation, keeping the same scale locks. The posture is ``no knobs'' at the derivation layer and sector--global changes only at the display layer; the framework is falsified by (a) disagreement between the two SI landings beyond stated uncertainty, or (b) failure of the sectorwise predictions beyond their single, globally declared policy bands.
\end{abstract}

\section*{Layer~0: Dimensionless invariants (fixed upstream)}
We assume as inputs only dimensionless identities:
\[
J(e^{t})=\cosh t-1,\qquad
\text{eight--tick minimal cycle},\qquad
\ln\varphi>0.
\]
These are the only dimensionless items the display layer may see.

\section*{Layer~1: Reality Bridge (semantics, falsifiable)}
We fix the bridge displays
\[
\frac{S}{\hbar}=J,\qquad
\frac{\tau_{\mathrm{rec}}}{\tau_{0}}=\frac{2\pi}{8\ln\varphi},\qquad
\frac{\lambda_{\mathrm{kin}}}{\ell_{0}}=\frac{2\pi}{8\ln\varphi},
\]
with $c=\ell_{0}/\tau_{0}$ and $\lambda_{\mathrm{kin}}=c\,\tau_{\mathrm{rec}}$. Two independent SI landings (time--first and length--first) must agree within declared combined uncertainty. This constitutes a \emph{single} pass/fail criterion; no regression, priors, or per--sector dials are allowed.

\section*{Layer~2: Bridge constants and anchor}
The constants used downstream are locked as follows:
\[
E_{\mathrm{coh}}=\varphi^{-5}\,\mathrm{eV},\qquad
\mu_\star=182.201~\mathrm{GeV}.
\]
Sector yardsticks use $A_B=B_B\,E_{\mathrm{coh}}\,\varphi^{r_0(B)}$ with $(B_B,r_0(B))$ frozen once per sector and then held fixed.

\section*{Layer~3: Discrete constructor and residue law at $\mu_\star$}
A finite constructor emits integers per species: a rung integer $r_i\in\mathbb Z$ and a word--charge $Z\in\mathbb Z$ (piecewise by charge/representation). The anchor residue is a closed form in $Z$:
\[
\mathcal F(Z)\;=\;\frac{\ln\!\bigl(1+Z/\varphi\bigr)}{\ln\varphi}.
\]
Equal--$Z$ classes have equal residues at the anchor. Rung integers are fixed (no fit knobs) and include a single global family torsion.

\section*{Layer~4: Standard--Model masses (fixed points, no per--species knobs)}
For each species in sector $B$,
\[
m_i\;=\;A_B\,\varphi^{\,r_i+f_B(m_i)}\qquad\text{with}\qquad f_B(m_i)\ \xrightarrow[\text{at }\mu_\star]{}\ \mathcal F(Z_i).
\]
Display policy is global per sector (e.g.\ QCD/QED loop orders, thresholds). When $Z_i=Z_j$,
\[
\left.\frac{m_i}{m_j}\right|_{\mu_\star}=\varphi^{\,r_i-r_j}\quad\text{(parameter--free anchor ratio)}.
\]
A single global change (e.g.\ scheme or input) moves the sector coherently; species--by--species adjustments are forbidden.

\section*{Layer~5: Information--limited gravity (galaxies $\rightarrow$ LSS)}
On linear scales in Newtonian gauge,
\[
k^{2}\Phi=4\pi G\,a^{2}\rho_{b}\,w(k,a)\,\delta_{b},\qquad
w(k,a)=1+\varphi^{-3/2}\,[a/(k\,\tau_{0})]^{\alpha},\quad
\alpha=\tfrac12\!\bigl(1-1/\varphi\bigr).
\]
The growth factor in the matter era integrates to
\[
D(a,k)=a\Bigl[\,1+\beta(k)\,a^{\alpha}\Bigr]^{1/(1+\alpha)},\qquad
\beta(k)=\tfrac{2}{3}\,\varphi^{-3/2}\,(k\,\tau_{0})^{-\alpha}.
\]
The same kernel used for galaxy phenomenology is carried into large--scale structure with no new parameters.

\section*{Predictions, invariants, and falsifiers}
\begin{itemize}
  \item \textbf{Bridge test (binary).} The time--first and length--first landings must agree within combined uncertainty at stated coverage. Failure falsifies the semantics.
  \item \textbf{Anchor invariants (spectra).} Equal--$Z$ residue degeneracy; exact anchor ratios $\bigl.m_i/m_j\bigr|_{\mu_\star}=\varphi^{r_i-r_j}$; closed--form residue $\mathcal F(Z)$.
  \item \textbf{Cosmology (linear).} Scale--dependent growth from $w(k,a)$; specific suppression in weak lensing at high multipoles; cluster counts without new matter components, under the same locks.
  \item \textbf{No per--species or per--galaxy rescue.} All shifts are sector--global; violations beyond the declared bands falsify the corresponding sector.
\end{itemize}

\section*{Scope and posture}
The derivation layer is parameter--free. Numerical displays follow from fixed semantics and standard constants. Sector predictions are auditable artifacts (one set of policies per sector) with exact anchor statements and narrow global bands. This is a framework for being ruled out cleanly.

% --- Optional engineering notes that do not introduce references ---
\section*{Engineering notes (non-normative)}
(1) Artifact discipline: ship the bridge calculator (both landings), the anchor residue audit, and the ILG growth kernel in meter--native units.
(2) Version locks and hashes should be part of the artifact manifest to enforce reproducibility.

% --- Clearly signaled unknowns if any ---
% Keep BLOCKER lines explicit and isolated.
% (Keep empty if you have no open dependencies for this capsule.)
% BLOCKER: none.

\end{document}
