\documentclass[11pt]{article}
\usepackage{amsmath,amssymb}

\title{Executive Summary:\\
\large Recognition Geometry: Inevitability Closure, Lawful Measurement Bridges, and Cross--Domain Audits}
\author{Jonathan Washburn\\
Recognition Science, Recognition Physics Institute\\
Austin, Texas, USA\\
\texttt{jon@recognitionphysics.org}}
\date{}

\begin{document}
\maketitle

\section*{Essence}
This paper presents a parameter--free architectural layer (\emph{Recognition Geometry}) and proves an \emph{Inevitability Closure}: once a short, lawful measurement bridge is in place, the architecture forces cross--domain consequences with no room for epicycles or parameter tuning. Numbers never enter the core; they appear only as finite certificates with interval errors that either pass or fail against the same invariants.

\section*{Core Claim}
\emph{Architecture} (eight--beat coverage/minimality with balanced $8$--windows, unique symmetric cost on the positive line with $J_{\log}(t)=\cosh t-1$, golden--ratio fixed point $\phi=1+\phi^{-1}$, and units identity $\ell_0/\tau_0=c$) $+$ \emph{Lawful Bridge} (units homomorphism, $\phi$--rung preservation, eight--beat invariants, and transport of Euler--Lagrange stationarity) $\Rightarrow$ \emph{unavoidable, testable outcomes across domains}.

\section*{What the Paper Proves}
\begin{itemize}
  \item \textbf{Mass ladder (parameter--free form).} Each species carries a rung $r\in\mathbb{Z}$ and fractional $f\in[0,1)$ with
  \[
     m \;=\; B\,E_{\mathrm{coh}}\,\phi^{\,r+f}, \qquad r\mapsto r+1 \;\Rightarrow\; m \mapsto \phi\,m,
  \]
  where $B>0$ is dimensionless and $E_{\mathrm{coh}}>0$ is a units anchor. This is enforced by the cost and eight--beat structure, not fitted.
  \item \textbf{Flat rotation kernel.} On intervals where enclosed mass is linear, $M_{\mathrm{enc}}(r)=\gamma r$, the circular speed is radius--independent:
  \[
     v(r)\equiv \sqrt{G\,\gamma}.
  \]
  Lawful bridges preserve this kernel.
  \item \textbf{Wedge analyticity via outer cancellation.} With $u$ a boundary datum on shifted half--planes, the outer $O=\exp(U+iV)$ (Poisson $U$ and conjugate $V$) cancels boundary phase and converts boundary control into interior energy. Under a Carleson budget compatible with $8$--window geometry, analytic quotients are Schur--bounded on admissible wedges and bounded at punctures, hence extend holomorphically (removable singularities).
  \item \textbf{No--cheat invariants.} No seven--tick full--parity coverage exists; no rung step other than multiplication by $\phi$ is definable under a lawful bridge. Attempts to ``rescue'' failures by reparametrization or units tweaks are excluded.
\end{itemize}

\section*{How It Works}
\textbf{Architecture} fixes the admissible increments (eight--beat minimality/balance) and the unique symmetric cost $J_{\log}(t)=\cosh t-1$ with stationary point at $t=0$. The \textbf{Lawful Bridge} preserves units ($\ell_0/\tau_0=c$), the $\phi$--rung law, eight--beat invariants, and Euler--Lagrange stationarity. The \textbf{Analytic Engines} provide: (i) outers on shifted half--planes with the correct boundary modulus, (ii) Carleson budgets adapted to Whitney geometry, and (iii) a stability window that supplies uniform bounds, together yielding wedge boundedness and removability.

\section*{Practical Outcomes}
\begin{itemize}
  \item \textbf{Audits, not fits.} Datasets are supplied as finite certificates (interval inequalities). They either pass the invariants or identify exactly which obligation fails.
  \item \textbf{Cross--domain leverage.} Passing in one domain tightens bounds in others (e.g., rung assignments constrain admissible mixing and flat zones).
  \item \textbf{Defensible simplicity.} The core is parameter--free; numbers live only in certificates. There is no room for ad hoc parameters to save a bad result.
\end{itemize}

\section*{Scope and Non--Goals}
The paper proves architectural necessity and wedge analyticity under budgets compatible with $8$--window geometry. It is not a numerical survey and introduces no fitted constants; empirical values (e.g., for $B$, $E_{\mathrm{coh}}$, budgets) belong to certificates outside the core text.

\section*{Team Acceptance Checklist (for submission)}
\begin{itemize}
  \item Architectural statements and the Inevitability Closure are presented parameter--free.
  \item Outer construction, budget bounds, and wedge removability are stated in the shifted half--plane setting.
  \item Certificate templates (mass, rotation, boundary) are included as examples with interval errors, not as data.
  \item Failure modes are explicit: non--$\phi$ rung, seven--tick coverage, or nonremovable wedge under admissible budgets.
\end{itemize}

\end{document}
