\documentclass[11pt]{article}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{geometry}
\geometry{margin=1in}

\title{Global Co-Identity Constraint (GCIC): Unity as a Gauge of Recognition Geometry}
\author{}
\date{}

\begin{document}

\maketitle

\section*{Statement of the Principle}
Let $L_0>0$ be the recognition anchor and $\varphi=\tfrac{1+\sqrt{5}}{2}$. Define the log-$\varphi$ coordinate
\[
x(\ell)\;:=\;\log_{\varphi}\!\big(\ell/L_0\big),\qquad \ell>0.
\]
\textbf{Global Co-Identity Constraint (GCIC).} There exists a single global phase $\Theta\in[0,1)$, defined modulo $1$, such that every pinned or steady recognition state satisfies
\[
x(\ell)\equiv \Theta \pmod{1}.
\]
Equivalently, the allowed recognition lengths occupy the $\varphi$-geometric ladder
\[
\ell_k\;=\;L_0\,\varphi^{\,k+\Theta},\qquad k\in\mathbb{Z},
\]
with a single, universe-wide $\Theta$.

\section*{Local Dynamics vs. Global Unity}
Local reproduction remains finite-rate: updates in $\ell$ propagate by re-creation at speed $c$. GCIC is not a signal; it is the definition of the steady-state manifold. Thus ``instantaneous unity'' is a gauge condition (global co-identity), not superluminal transport.

\section*{Recognition Cost and Stationary Set}
A minimal log-periodic recognition cost achieving this constraint is
\[
C(\ell;\Theta)\;=\;C_\star+\kappa\!\left[1-\cos\!\big(2\pi\,[x(\ell)-\Theta]\big)\right],\qquad \kappa>0,
\]
with stationary minima at $\ell=\ell_k$ for all $k\in\mathbb{Z}$ and positive local curvature
\[
\frac{d^2 C}{d\ell^2}\Big|_{\ell=\ell_k}
=\frac{(2\pi)^2\kappa}{\ell_k^{\,2}(\ln\varphi)^2}\;>\;0.
\]
The two-phase alternation is the sign of the logarithmic slope about a minimum:
\[
\text{contractive }(+)\;\Longleftrightarrow\; -\,\partial_\ell C>0,\qquad
\text{expansive }(-)\;\Longleftrightarrow\; -\,\partial_\ell C<0.
\]

\section*{Reciprocity and Scale-Action}
Let the nine ladder slots be $n\in\{-4,-3,\dots,+3,+4\}$ with
\[
\ell(n)\;=\;L_0\,\varphi^{-n},\qquad \nu(n)\;=\;\frac{c}{\ell(n)}=\nu_0\,\varphi^{\,n},\quad \nu_0=\frac{c}{L_0}.
\]
GCIC enforces the global reciprocity constraint
\[
\ell(+n)\,\ell(-n)\;=\;L_0^2,
\]
which is a conserved \emph{scale-action} across the ledger. This invariance is independent of the value of $\Theta$.

\section*{Testable Consequences}
\textbf{(i) $\varphi$-band clustering on a log-axis.} Any phenomenon that ``snaps'' to recognition minima exhibits geometric banding with ratio $\varphi$ when plotted in $x(\ell)$; paired (contractive/expansive) features mirror about the central node.\\
\textbf{(ii) Loop quantization.} For any closed loop of coupled cavities or waveguides, the sum of log-$\varphi$ phase shifts along the loop is an integer: $\sum \Delta x \in \mathbb{Z}$.\\
\textbf{(iii) Unity under rewiring.} If $\Theta$ is held, the pinned pattern in a $\varphi$-ladder network is invariant under graph rewiring that preserves total loop phase; injection point does not change the resting set $\{\ell_k\}$.

\end{document}