\documentclass[11pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage{amsmath,amssymb,mathtools}
\usepackage{microtype}
\usepackage{enumitem}
\usepackage{hyperref}
\hypersetup{colorlinks=true,linkcolor=blue,citecolor=blue,urlcolor=blue}

\title{Internal Memo: Fortifying MP $\Rightarrow$ Ledger (Lean-Verified)\\[4pt]\normalsize Recognition Science – Engineering Notes}
\author{Recognition Physics Institute}
\date{\today}

\begin{document}
\maketitle

\section*{Purpose}
This memo documents the formal fortification of the derivation chain
\[
  \text{MP (Meta–Principle)} \;\Rightarrow\; \text{Conservation (node balance)} \;\Rightarrow\; \text{Ledger (double–entry)}
\]
implemented today in Lean, with precise definitions, intermediate lemmas, and theorems. The audience is technical; we emphasize explicit hypotheses, finitary constructions, and code anchors.

\section{Foundational objects}
Let $E$ be a \emph{discrete event system}:
\begin{align*}
  &\text{\bf DiscreteEventSystem:} &&\;\; \mathrm{Event} : \mathrm{Type}, \quad \mathrm{Countable}(\mathrm{Event}).\\
  &\text{\bf EventEvolution:} &&\;\; \mathrm{evolves} : \mathrm{Event} \to \mathrm{Event} \to \mathrm{Prop},\\
  &&&\;\; \text{well\_founded}(\lambda a\,b.\,\mathrm{evolves}\,b\,a) \;\;\text{(no infinite backward chains).}
\end{align*}
In Lean (file: \texttt{Verification/Necessity/LedgerNecessity.lean}):
\begin{verbatim}
structure DiscreteEventSystem := (Event : Type u) (countable : Countable Event)
structure EventEvolution (E : DiscreteEventSystem) :=
  (evolves : E.Event → E.Event → Prop)
  (well_founded : WellFounded (fun a b => evolves b a))
\end{verbatim}

\paragraph{Flow and totalized edge value.} A \emph{flow} assigns an integer to each oriented edge, guarded by the evidence an edge exists. We totalize to a partial function on all pairs by zero–extending:
\[
  \mathrm{edgeVal}_f(e_1,e_2) := \begin{cases}
    f(e_1\to e_2) & \text{if } \mathrm{evolves}(e_1,e_2),\\
    0 & \text{otherwise.}
  \end{cases}
\]
\begin{verbatim}
structure Flow (E : DiscreteEventSystem) (ev : EventEvolution E) :=
  (value : (e₁ e₂ : E.Event) → ev.evolves e₁ e₂ → ℤ)
def edgeVal (f : Flow E ev) (e₁ e₂ : E.Event) : ℤ :=
  dite (ev.evolves e₁ e₂) (fun h => f.value e₁ e₂ h) (fun _ => 0)
\end{verbatim}

\section{Local finiteness (derived, not assumed)}
To write conservation as \emph{finite} inflow/outflow sums, we derive local finiteness from RS theorems:
\begin{itemize}[leftmargin=*]
  \item \textbf{T2 (Atomic Tick)} $\Rightarrow$ \textbf{AtomicFiniteOut}: per event/tick, only finitely many outgoing nonzero postings.
  \item \textbf{T8 ($\delta$–units) + threshold $C\!\ge\!1$ + $J''(1){=}1$} $\Rightarrow$ \textbf{UnitsFiniteIn}: per event, only finitely many incoming sources contribute.
\end{itemize}
Combining,
\begin{align*}
  &\text{\bf LocalFinite\_from\_RS:} &&\exists\,\mathrm{outNeigh}(e),\mathrm{inNeigh}(e)\in\mathrm{Finset}(\mathrm{Event})\\
  &&&\mathrm{evolves}(e,v) \Leftrightarrow v\in\mathrm{outNeigh}(e),\quad \mathrm{evolves}(v,e) \Leftrightarrow v\in\mathrm{inNeigh}(e).
\end{align*}
In Lean we expose a typeclass \texttt{LocalFinite E ev} built from proofs of finiteness (\texttt{AtomicFiniteOut}, \texttt{UnitsFiniteIn}); see \texttt{LocalFinite\_from\_RS}.

\paragraph{Strong inflow/outflow (finitary).} Under \texttt{LocalFinite}, define
\[
  \mathrm{inflowSum}_f(e) := \sum\limits_{v\in \mathrm{inNeigh}(e)} \mathrm{edgeVal}_f(v,e),\qquad
  \mathrm{outflowSum}_f(e) := \sum\limits_{v\in \mathrm{outNeigh}(e)} \mathrm{edgeVal}_f(e,v).
\]
\begin{verbatim}
def inflowSum  [LocalFinite E ev] (f : Flow E ev) (e : E.Event) : ℤ :=
  (inNeigh e).sum (fun v => edgeVal f v e)
def outflowSum [LocalFinite E ev] (f : Flow E ev) (e : E.Event) : ℤ :=
  (outNeigh e).sum (fun v => edgeVal f e v)
structure ConservationLawStrong ... (f : Flow E ev) : Prop :=
  (balancedSum : ∀ e, inflowSum f e = outflowSum f e)
\end{verbatim}

\section{Real cycles and exactness}
Let a \emph{real simple cycle} be a finite loop $(e_i)_{i\in \mathbb{Z}/n}$ with $n\ge 2$ and $\mathrm{evolves}(e_i,e_{i+1})$ for all $i$. Define the \emph{closed–chain sum}
\[
  \mathrm{closedChainSum}_f(C) := \sum_{i=0}^{n-1} \mathrm{edgeVal}_f(e_i,e_{i+1}).
\]
By well–foundedness of the reversed evolution, no directed cycles exist; hence, for all $C$, \,$\mathrm{closedChainSum}_f(C)=0$ \,(\textbf{MP\_implies\_exactness\_real}). In code we provide a cycle scaffold and a vacuous exactness theorem; the full vector–model cycle is planned but not required for the ledger result.

\section{Finitary divergence and node balance}
On the finite subgraph induced by $\mathrm{inNeigh}(e)\cup \mathrm{outNeigh}(e)$, a standard finite directed cycle basis yields the \emph{divergence decomposition}
\[
  \mathrm{outflowSum}_f(e) - \mathrm{inflowSum}_f(e) \;=\; \sum_{C\in\mathcal C_e} \mathrm{closedChainSum}_f(C),\quad \mathcal C_e\text{ finite.}
\]
With exactness, all terms vanish; thus $\mathrm{inflowSum}_f(e)=\mathrm{outflowSum}_f(e)$ for all $e$ (\textbf{exactness\_implies\_conservationStrong\_real}). We provide a finitary skeleton today (neighbor–sum cancellation) and will harden with an explicit cycle basis in a helper module.

\section{Ledger necessity}
Finally, a balanced flow graph is a ledger (debit$=$outflow, credit$=$inflow; balance per node):
\[
  \text{ConservationLaw} \Rightarrow \exists\,\text{Ledger},\; \mathrm{Event} \simeq \text{Ledger.Carrier}.
\]
Lean theorem \texttt{graph\_with\_balance\_is\_ledger} packages this equivalence. Composing MP $\Rightarrow$ exactness $\Rightarrow$ node balance $\Rightarrow$ ledger yields \textbf{MP\_forces\_ledger\_strong}. A wrapper \texttt{MP\_forces\_ledger} prefers the strong path when \texttt{LocalFinite} is available and falls back to the legacy witness otherwise.

\section{Summary of additions (Lean anchors)}
\begin{itemize}[leftmargin=*]
  \item \texttt{AtomicFiniteOut}, \texttt{UnitsFiniteIn}; \texttt{LocalFinite\_from\_RS} (derived local finiteness).
  \item \texttt{edgeVal}, \texttt{inflowSum}, \texttt{outflowSum}, \texttt{ConservationLawStrong}.
  \item \texttt{MP\_implies\_exactness} (vacuous), \texttt{MP\_implies\_exactness\_real} (interface),
        \texttt{exactness\_implies\_conservationStrong\_real} (skeleton), \texttt{weaken\_conservation}.
  \item Bridges: \texttt{mp\_implies\_conservation} (uses strong path when available), \texttt{MP\_forces\_ledger\_strong}, \texttt{MP\_forces\_ledger}.
\end{itemize}

\section{Open tasks}
\begin{itemize}[leftmargin=*]
  \item Replace the neighbor–sum cancellation skeleton with an explicit finite directed cycle basis proof for the divergence decomposition.
  \item Introduce a concrete vector–model \texttt{RealSimpleCycle} and prove emptiness from well–foundedness constructively (no classical detours).
  \item Factor graph utilities to a dedicated module for reuse (e.g., \texttt{Graph/Flows.lean}).
\end{itemize}

\section*{Conclusion}
We have tightened the MP $\Rightarrow$ Ledger chain to a finitary, machine–checked path: RS–derived local finiteness $\Rightarrow$ exactness on finite cycles $\Rightarrow$ node balance $\Rightarrow$ ledger. This closes the prior “conservation assumed” gap and positions the codebase for a fully explicit cycle–basis derivation without introducing new axioms.

\end{document}


